A Memory-Bounded, Deterministic and Terminating Semantics for the Synchronous Programming Language~\CEU

### 2018-04-15 02-ceu.tex
\small no listings da Seção 2 (ao invés de \footnotesize)
Usar \footnotesize nesses listings não ganha quase nada de espaço e deixa o
código mais difícil de ler.  Também tirei o \small do \code, para ficar
uniforme com a Seção 3.  O ganho do texttt-small é muito pequeno, e fica
feio quando misturado com matemática (como na seção 3).

\bgroup
\def\T<#1>{\langle\mathit{#1}\rangle}
\def\C#1{\rmfamily\itshape//~#1}
\def\CC#1#2{\hfill\makebox[#1][l]{\C{#2}}}
\begin{lstlisting}[
  basicstyle=\ttfamily\footnotesize,
  caption={The concrete syntax of \CEU.},
  label={lst.syntax},
]
|\C{Declarations}|
input $\T<type>$ $\T<id>$; | \CC{19.5em}{declares an external input event} |
event $\T<type>$ $\T<id>$; | \CC{19.5em}{declares an internal event}       |
var   $\T<type>$ $\T<id>$; | \CC{19.5em}{declares a variable}              |

|\C{Event handling}|
$\T<id>$ = await $\T<id>$; | \CC{19.5em}{awaits an event and assigns the received value} |
emit $\T<id>$($\T<exp>$); | \CC{19.5em}{emits an event passing a value} |

|\C{Control flow}|
$\T<stmt>$ ; $\T<stmt>$ | \CC{10em}{sequence}     |
if $\T<exp>$ then $\T<stmts>$ else $\T<stmts>$ end | \CC{10em}{conditional}  |
loop do $\T<stmts>$ end | \CC{10em}{repetition}   |
every $\T<id>$ in $\T<id>$ do $\T<stmts>$ end | \CC{10em}{event iteration}   |
finalize [$\T<stmts>$] with $\T<stmts>$ end | \CC{10em}{finalization} |

|\C{Logical parallelism}|
par/or  do $\T<stmts>$ with $\T<stmts>$ end | \CC{13em}{aborts when any side ends}      |
par/and do $\T<stmts>$ with $\T<stmts>$ end | \CC{13em}{terminates when all sides end} |

|\C{Assignment \& Integration with C}|
$\T<id>$ = $\T<exp>$; | \CC{19.5em}{assigns a value to a variable}              |
_$\T<id>$($\T<exps>$) | \CC{19.5em}{calls a C function (id starts with `\_'\,)} |
\end{lstlisting}
\egroup

@@

\begin{lstlisting}[
  numbers=left,
  basicstyle=\ttfamily\footnotesize,
  float=h,
  caption={A \CEU program that toggles a LED on radio receive,
           terminating on a button press always with the LED~off.},
  label={lst.ceu},
]
input void BUTTON;
input void RADIO_RECV;
par/or do
    await BUTTON;
with
    finalize with
        _led(0);
    end
    loop do
        _led(1);
        await RADIO_RECV;
        _led(0);
        await RADIO_RECV;
    end
end
\end{lstlisting}

@@

\CEU defines time as a discrete sequence of reactions to unique external
input events received from the environment.
%
Each input event delimits a new logical unit of time that triggers an
associated reaction.
%
The life-cycle of a program in \CEU can be summarized as
follows~\cite{ceu.sensys13}:
%
\begin{enumerate}[i]
\item The program initiates a ``boot reaction'' in a single trail (parallel
      constructs may create new trails).
\item Active trails execute until they await or terminate, one after
      another.  This step is called a \emph{reaction chain}, and always runs in
      bounded time.
\item When all trails are blocked, the program goes idle and the environment
      takes control.
\item On the occurrence of a new external input event, the environment
      awakes \emph{all} trails awaiting that event, and the program goes back to
      step~(ii).
\end{enumerate}

@@

The sequential processing of external input events induces a discrete notion of
time in \CEU, as illustrated in Figure~\ref{fig.ticks}.
%
The continuous timeline shows an absolute reference clock with ``physical
timestamps'' for the event occurrences (e.g., event~\code{C} occurs at
$17ms521us$).
%
The discrete timeline shows how the same occurring events fit in the logical
notion of time of \CEU.
%
The boot reaction \code{boot-0} happens before any input, at program startup.
%
Event~\code{A} ``physically'' occurs during \code{boot-0} but, because time
is discrete, its corresponding reaction only executes afterwards, at logical
instant~\code{A-1}.
%
Similarly, event~\code{B} occurs during~\code{A-1} and its reaction is
postponed to execute at~\code{B-2}.
%
Event~\code{C} also occurs during~\code{A-1} but its reaction must also wait
for~\code{B-2} to execute and so it is postponed to execute at~\code{C-3}.
%
Finally, event~\code{D} occurs during an idle period and can start immediately
at \code{D-4}.
%
%Finally, two instances of event~\code{E} occur during~\code{D-4}; they are
%handled in the subsequent reactions~\code{E-5} and~\code{E-6}.

\begingroup
\begin{figure}[t]
\centering
\begin{tikzpicture}[x=.9em,y=.9em,
  font=\footnotesize\sffamily,>=stealth,
  time/.style={font=\scriptsize\sffamily},
  cross/.style={inner sep=0pt,outer sep=0pt},
  evt/.style={fill=white,inner sep=1.85pt,outer sep=0pt},
  rng/.style={fill=white,inner sep=1pt,outer sep=0pt,
    font=\ttfamily\scriptsize}]
  \draw[->]
  (0,0)
    node[align=right,anchor=east,xshift=-.5em]
        {Continuous\\Timeline\\(Real World)}
    node[anchor=north west,xshift=-.5em,time]
        {0ms000us}
    node[anchor=north west,xshift=8.75em,time]
        {17ms521us}
  -- (18.5,0)
    node[anchor=north east,xshift=.1em,time]
        {30ms000us};
  \node[cross] at (0,0){x};
  %%
  \coordinate(Ext) at ($(0,0)!.5!(0,-5)$);
  \coordinate(A) at (Ext-|4,0);
  \coordinate(B) at (Ext-|5.5,0);
  \coordinate(C) at (Ext-|9,0);
  \coordinate(D) at (Ext-|15,0);
  %%
  \draw[semithick,dotted] (0,1-|A) -- (0,-6-|A);
  \draw[semithick,dotted] (0,1-|B) -- (0,-6-|B);
  \draw[semithick,dotted] (0,1-|C) -- (0,-6-|C);
  \draw[semithick,dotted] (0,1-|D) -- (0,-5-|D);
  \node[evt](Aevt) at (A) {\textbf{A}};
  \node[evt](Bevt) at (B) {\textbf{B}};
  \node[evt](Cevt) at (C) {\textbf{C}};
  \node[evt](Devt) at (D) {\textbf{D}};
  \node[anchor=east,xshift=-.5em] at (A){\emph{External Input}};
  %%
  \draw[-]
  (0,-5)
    node[align=right,anchor=east,xshift=-.5em]
        {Discrete\\Timeline\\(\CEU)}
  -- (18.5,-5);
  \node[cross](P0) at (0,-5){x};
  \node[cross](P1) at (4.75,-5){x};
  \node[cross](P2) at (10,-5){x};
  \node[cross](P3) at (12.5,-5){x};
  \node[cross](P4) at (14.5,-5){x};
  \node[cross](P5) at (15.5,-5){x};
  \node[cross](P6) at (18.5,-5){x};
  \node[rng] at ($(P0)!.5!(P1)$){boot-0};
  \node[rng] at ($(P1)!.5!(P2)$){A-1};
  \node[rng] at ($(P2)!.5!(P3)$){B-2};
  \node[rng] at ($(P3)!.5!(P4)$){C-3};
  \node[rng] at ($(P5)!.5!(P6)$){D-4};
  \draw[decorate,decoration={brace,raise=-1pt,mirror,amplitude=2pt}]
  ($(P4)+(0,-.5)$) -- node[below,time] {Idle period} ($(P5)+(0,-.5)$);
  %%
  \begin{scope}[shorten >=1.5pt]
  \draw[->] (Aevt) -- (P1);
  \draw[->] (Bevt) -- (P2);
  \draw[->] (Cevt) -- (P3);
  \draw[->] (Devt) -- (P5);
  \end{scope}
\end{tikzpicture}
%\includegraphics[width=\columnwidth]{tick_min}
\caption{The discrete notion of time in \CEU.}
\label{fig.ticks}
\end{figure}
\endgroup

@@

\begin{lstlisting}[
  numbers=left,
  basicstyle=\ttfamily\footnotesize,
  float=h,
  caption={A ``subroutine'' that increments its argument.},
  label={lst.sub},
]
event int* inc;         // declares subroutine "inc"
par/or do
    var int* p;
    every p in inc do   // implements "inc" through an event iterator
        *p = *p + 1;
    end
with
    var int v = 1;
    emit inc(&v);       // calls "inc"
    emit inc(&v);       // calls "inc"
    _assert(v==3);      // asserts result after the two returns
end
\end{lstlisting}
%\vskip-\baselineskip

@@

At first sight, the constructions ``\code{every e do <...> end}'' and
``\code{loop do await e; <...> end}'' seem to be equivalent.
However, the \code{loop} variation would not compile since it does not contain
an external input \code{await} (\code{e} is an internal event).
%
The \code{every} variation compiles because event iterators have an additional
syntactic restriction that they cannot contain \code{break} or \code{await}
statements.
This restriction guarantees that an iterator never terminates from itself and,
thus, always awaits its identifying event, essentially behaving as a safe
blocking point in the program.
%
For this reason, the restriction that execution paths within loops must
contain at least one external \code{await} is extended to alternatively contain
an \code{every} statement.

@@

\begingroup
\begin{figure}[t]
\begin{minipage}[h]{0.45\linewidth}
\begin{lstlisting}[numbers=right,basicstyle=\ttfamily\footnotesize]
input void A;
input void B;
var int x = 1;
par/and do
    await A;
    x = x + 1;
with
    await B;
    x = x * 2;
end
\end{lstlisting}
\small{\ax Accesses to \code{x} are never\\\phantom{\ax}\relax concurrent.}
\end{minipage}%
%
\begin{minipage}[h]{0.53\linewidth}
%\begin{lstlisting}
\begin{lstlisting}[xleftmargin=2em,basicstyle=\ttfamily\footnotesize]
input void A;
// (empty line)
var int y = 1;
par/and do
    await A;
    y = y + 1;
with
    await A;
    y = y * 2;
end

\end{lstlisting}
\small{\bx Accesses to \code{y} are concurrent\\
\phantom{\bx}\relax but still deterministic.}
\end{minipage}%
%\rule{8.4cm}{0.37pt}
\caption{Shared-memory concurrency in \CEU:
\ax is safe because the trails access \code{x} atomically in different
reactions;
\bx is unsafe because both trails access \code{y} in the same reaction}
\label{lst.shared}
\end{figure}
\endgroup

@@

\CEU also enforces, at compile time, the use of \code{finalize} for system
calls that deal with pointers representing resources, as illustrated in the two
examples of Figure~\ref{lst.fin.ceu}.
%
%\begin{itemize}
 If \CEU \textbf{passes} a pointer to a system call (ln.~\ax:5), the
          pointer represents a \textbf{local} resource (ln.~\ax:2) that
          requires finalization (ln.~\ax:7).
 If \CEU \textbf{receives} a pointer from a system call return
          (ln.~ \bx:4), the pointer represents an \textbf{external} resource
          (ln.~\bx:2) that requires finalization (ln.~\bx:6).
%\end{itemize}

@@

\begin{figure}[t]
\begin{minipage}[t]{0.45\linewidth}
\begin{lstlisting}[numbers=right,basicstyle=\ttfamily\footnotesize]
par/or do
   var _msg_t msg;
   <...> // prepare msg
   finalize
      _send(&msg);
   with
      _cancel(&msg);
   end
   await SEND_ACK;
with
   <...>
end
//
\end{lstlisting}
\small{\ax\!Local resource finalization\rlap{.}}
\end{minipage}%
%
\begin{minipage}[t]{0.53\linewidth}
\begin{lstlisting}[xleftmargin=2em,basicstyle=\ttfamily\footnotesize]
par/or do
   var _FILE* f;
   finalize
      f = _fopen(...);
   with
      _fclose(f);
   end
   _fwrite(..., f);
   await A;
   _fwrite(..., f);
with
   <...>
end
\end{lstlisting}
\small{\quad\bx\!External resource finalization\rlap{.}}
\end{minipage}%
%\rule{8.4cm}{0.37pt}
\caption{%
\CEU enforces the use of finalization to prevent dangling pointers and memory leaks.}
\label{lst.fin.ceu}
\end{figure}

>>>>

\def\C#1{\rmfamily\itshape//~#1}
\def\CC#1#2{\hfill\makebox[#1][l]{\C{#2}}}
\def\T<#1>{\langle\mathit{#1}\rangle}
\bgroup
\begin{lstlisting}[
  basicstyle=\ttfamily\small,
  caption={The concrete syntax of \CEU.},
  label={lst.syntax},
]
|\C{Declarations}|
input $\T<type>$ $\T<id>$; |\CC{19.5em}{declares an external input event} |
event $\T<type>$ $\T<id>$; |\CC{19.5em}{declares an internal event}       |
var   $\T<type>$ $\T<id>$; |\CC{19.5em}{declares a variable}              |

|\C{Event handling}|
$\T<id>$ = await $\T<id>$; |\CC{19.5em}{awaits event and assigns the received value} |
emit $\T<id>$($\T<expr>$); |\CC{19.5em}{emits event passing a value} |

|\C{Control flow}|
$\T<stmt>$ ; $\T<stmt>$ | \CC{8em}{sequence}     |
if $\T<expr>$ then $\T<stmt>$ else $\T<stmt>$ end | \CC{8em}{conditional}  |
loop do $\T<stmt>$ end | \CC{8em}{repetition}   |
every $\T<id>$ in $\T<id>$ do $\T<stmt>$ end | \CC{8em}{event iteration}   |
finalize [$\T<stmt>$] with $\T<stmt>$ end | \CC{8em}{finalization} |

|\C{Logical parallelism}|
par/or  do $\!\T<stmt>\!$ with $\!\T<stmt>\!$ end |\CC{13.5em}{aborts if any side ends}      |
par/and do $\!\T<stmt>\!$ with $\!\T<stmt>\!$ end |\CC{13.5em}{terminates if all sides end} |

|\C{Assignment \& Integration with C}|
$\T<id>$ = $\T<expr>$; | \CC{19.5em}{assigns a value to a variable}              |
_$\T<id>$($\T<exprs>$) | \CC{19.5em}{calls a C function (id starts with `\_'\,)} |
\end{lstlisting}
\egroup

@@

\begin{lstlisting}[
  numbers=left,
  basicstyle=\ttfamily\small,
  float=h,
  caption={A \CEU program that toggles a LED on radio receive,
           terminating on a button press always with the LED~off.},
  label={lst.ceu},
]
input void BUTTON;
input void RADIO_RECV;
par/or do
    await BUTTON;
with
    finalize with
        _led(0);
    end
    loop do
        _led(1);
        await RADIO_RECV;
        _led(0);
        await RADIO_RECV;
    end
end
\end{lstlisting}

@@

\CEU defines time as a discrete sequence of reactions to unique external
input events received from the environment.
%
Each input event delimits a new logical unit of time that triggers an
associated reaction.
%
The life-cycle of a program in \CEU can be summarized as
follows~\cite{ceu.sensys13}:
%
\begin{enumerate}[i.]
\item The program initiates a ``boot reaction'' in a single trail (parallel
      constructs may create new trails).
\item Active trails execute until they await or terminate, one after
      another.  This step is called a \emph{reaction chain}, and always runs in
      bounded time.
\item When all trails are blocked, the program goes idle and the environment
      takes control.
\item On the occurrence of a new external input event, the environment
      awakes \emph{all} trails awaiting that event, and the program goes back to
      step~(ii).
\end{enumerate}

@@

The sequential processing of external input events induces a discrete notion of
time in \CEU, as illustrated in Figure~\ref{fig.ticks}.
%
The continuous timeline shows an absolute reference clock with ``physical
timestamps'' for the event occurrences (e.g., event~\code{C} occurs at
17ms521us).
%
The discrete timeline shows how the same occurring events fit in the logical
notion of time of \CEU.
%
The boot reaction \code{boot-0} happens before any input, at program startup.
%
Event~\code{A} ``physically'' occurs during \code{boot-0} but, because time
is discrete, its corresponding reaction only executes afterwards, at logical
instant~\code{A-1}.
%
Similarly, event~\code{B} occurs during~\code{A-1} and its reaction is
postponed to execute at~\code{B-2}.
%
Event~\code{C} also occurs during~\code{A-1} but its reaction must also wait
for~\code{B-2} to execute and so it is postponed to execute at~\code{C-3}.
%
Finally, event~\code{D} occurs during an idle period and can start immediately
at \code{D-4}.
%
%Finally, two instances of event~\code{E} occur during~\code{D-4}; they are
%handled in the subsequent reactions~\code{E-5} and~\code{E-6}.

\begingroup
\begin{figure}[h]
\centering
\begin{tikzpicture}[x=1em,y=1em,
  font=\footnotesize\sffamily,>=stealth,
  time/.style={font=\scriptsize\sffamily},
  cross/.style={inner sep=0pt,outer sep=0pt},
  evt/.style={fill=white,inner sep=1.85pt,outer sep=0pt},
  rng/.style={fill=white,inner sep=1pt,outer sep=0pt,
    font=\ttfamily\footnotesize}]
  \draw[->]
  (0,0)
    node[align=right,anchor=east,xshift=-.5em]
        {Continuous\\Timeline\\(Real World)}
    node[anchor=north west,xshift=-.5em,time]
        {0ms000us}
    node[anchor=north west,xshift=8.75em,time]
        {17ms521us}
  -- (18.5,0)
    node[anchor=north east,xshift=.1em,time]
        {30ms000us};
  \node[cross] at (0,0){x};
  %%
  \coordinate(Ext) at ($(0,0)!.5!(0,-5)$);
  \coordinate(A) at (Ext-|4,0);
  \coordinate(B) at (Ext-|5.5,0);
  \coordinate(C) at (Ext-|9,0);
  \coordinate(D) at (Ext-|15,0);
  %%
  \draw[semithick,dotted] (0,1-|A) -- (0,-6-|A);
  \draw[semithick,dotted] (0,1-|B) -- (0,-6-|B);
  \draw[semithick,dotted] (0,1-|C) -- (0,-6-|C);
  \draw[semithick,dotted] (0,1-|D) -- (0,-5-|D);
  \node[evt](Aevt) at (A) {\texttt{\textbf{A}}};
  \node[evt](Bevt) at (B) {\texttt{\textbf{B}}};
  \node[evt](Cevt) at (C) {\texttt{\textbf{C}}};
  \node[evt](Devt) at (D) {\texttt{\textbf{D}}};
  \node[anchor=east,xshift=-.5em] at (A){\emph{External Input}};
  %%
  \draw[-]
  (0,-5)
    node[align=right,anchor=east,xshift=-.5em]
        {Discrete\\Timeline\\(\CEU)}
  -- (18.5,-5);
  \node[cross](P0) at (0,-5){x};
  \node[cross](P1) at (4.75,-5){x};
  \node[cross](P2) at (10,-5){x};
  \node[cross](P3) at (12.5,-5){x};
  \node[cross](P4) at (14.5,-5){x};
  \node[cross](P5) at (15.5,-5){x};
  \node[cross](P6) at (18.5,-5){x};
  \node[rng] at ($(P0)!.5!(P1)$){boot-0};
  \node[rng] at ($(P1)!.5!(P2)$){A-1};
  \node[rng] at ($(P2)!.5!(P3)$){B-2};
  \node[rng] at ($(P3)!.5!(P4)$){C-3};
  \node[rng] at ($(P5)!.5!(P6)$){D-4};
  \draw[decorate,decoration={brace,raise=-1pt,mirror,amplitude=2pt}]
  ($(P4)+(0,-.5)$) -- node[below,time] {Idle period} ($(P5)+(0,-.5)$);
  %%
  \begin{scope}[shorten >=1.5pt]
  \draw[->] (Aevt) -- (P1);
  \draw[->] (Bevt) -- (P2);
  \draw[->] (Cevt) -- (P3);
  \draw[->] (Devt) -- (P5);
  \end{scope}
\end{tikzpicture}
\vskip-.6em
%\includegraphics[width=\columnwidth]{tick_min}
\caption{The discrete notion of time in \CEU.}
\label{fig.ticks}
\end{figure}
\endgroup

@@

\begin{lstlisting}[
  numbers=left,
  basicstyle=\ttfamily\small,
  float=h,
  caption={A ``subroutine'' that increments its argument.},
  label={lst.sub},
]
event int* inc;       |\C{declares subroutine "inc"}|
par/or do
    var int* p;
    every p in inc do |\C{implements "inc" through event iterator}|
        *p = *p + 1;
    end
with
    var int v = 1;
    emit inc(&v);     |\C{calls "inc"}|
    emit inc(&v);     |\C{calls "inc"}|
    _assert(v==3);    |\C{asserts result after the two returns}|
end
\end{lstlisting}
%\vskip-\baselineskip

@@

At first sight, the constructions ``\code{every e do $\T<\cdots>$ end}'' and
``\code{loop do await e; $\T<\cdots>$ end}'' seem to be equivalent.
However, the \code{loop} variation would not compile since it does not contain
an external input \code{await} (\code{e} is an internal event).
%
The \code{every} variation compiles because event iterators have an additional
syntactic restriction that they cannot contain \code{break} or \code{await}
statements.
This restriction guarantees that an iterator never terminates from itself and,
thus, always awaits its identifying event, essentially behaving as a safe
blocking point in the program.
%
For this reason, the restriction that execution paths within loops must
contain at least one external \code{await} is extended to alternatively contain
an \code{every} statement.

@@

\begingroup
\begin{figure}[t]
\begin{minipage}[t]{0.45\linewidth}
\begin{lstlisting}[numbers=right,basicstyle=\ttfamily\small]
input void A;
input void B;
var int x = 1;
par/and do
    await A;
    x = x + 1;
with
    await B;
    x = x * 2;
end
\end{lstlisting}
\small{\ax\ Accesses to \code{x} are never concurrent.}
\end{minipage}%
%
\begin{minipage}[t]{0.53\linewidth}
%\begin{lstlisting}
\begin{lstlisting}[xleftmargin=1.75em,basicstyle=\ttfamily\small]
input void A;
|\C{(empty line)}|
var int y = 1;
par/and do
    await A;
    y = y + 1;
with
    await A;
    y = y * 2;
end

\end{lstlisting}
\hskip1.5em
\begin{minipage}[t]{.9\columnwidth}
\small{\bx\ Accesses to \code{y} are concurrent but still deterministic.}
\end{minipage}%
\end{minipage}%
%\rule{8.4cm}{0.37pt}
\caption{Shared-memory concurrency in \CEU:
\ax\ is safe because the trails access \code{x} atomically in different
reactions;
\bx\ is unsafe because both trails access \code{y} in the same reaction}
\label{lst.shared}
\end{figure}
\endgroup

@@

\CEU also enforces, at compile time, the use of \code{finalize} for system
calls that deal with pointers representing resources, as illustrated in the two
examples of Figure~\ref{lst.fin.ceu}.
%
%\begin{itemize}
 If \CEU \emph{passes} a pointer to a system call (ln.~\ax:5), the
          pointer represents a \emph{local} resource (ln.~\ax:2) that
          requires finalization (ln.~\ax:7).
 If \CEU \emph{receives} a pointer from a system call return
          (ln.~ \bx:4), the pointer represents an \emph{external} resource
          (ln.~\bx:2) that requires finalization (ln.~\bx:6).
%\end{itemize}

@@

\begin{figure}[t]
\begin{minipage}[t]{0.45\linewidth}
\begin{lstlisting}[numbers=right,basicstyle=\ttfamily\small]
par/or do
   var _msg_t msg;
   $\T<\cdots>$   |\C{prepare msg}|
   finalize
      _send(&msg);
   with
      _cancel(&msg);
   end
   await SEND_ACK;
with
   $\T<\cdots>$
end
|\C{(empty line)}|
\end{lstlisting}
\small{\ax\ Local resource finalization.}
\end{minipage}%
%
\begin{minipage}[t]{0.53\linewidth}
\begin{lstlisting}[xleftmargin=1.75em,basicstyle=\ttfamily\small]
par/or do
   var _FILE* f;
   finalize
      f = _fopen($\dots$);
   with
      _fclose(f);
   end
   _fwrite($\dots$, f);
   await A;
   _fwrite($\dots$, f);
with
   $\T<\cdots>$
end
\end{lstlisting}
\hskip1em
\begin{minipage}[t]{.94\columnwidth}
\small{\bx\ External resource finalization.}
\end{minipage}%
\end{minipage}%
%\rule{8.4cm}{0.37pt}
\caption{%
\CEU enforces the use of finalization to prevent dangling pointers and memory leaks.}
\label{lst.fin.ceu}
\end{figure}
